High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales

### 2021-07-23 inst/manuscript/manuscript.tex
Added dataset documentation, manuscript statement (EarthArxiv), several datasets

%\begin{center}
This pre-print not yet undergone peer review. It has been submitted to
the \emph{Frontiers in Artificial Intelligence}. This version will be
updated as it is revised, and the final published version will be
accessible through its DOI link.
%\end{center}

@@

\begin{abstract}
Understanding agricultural change requires reliable, frequently updated
maps that describe the characteristics of croplands. Such data are often
unavailable for regions dominated by smallholder agricultural systems,
which are particularly challenging for remote sensing. To overcome these
challenges, we designed a system to minimize several sources of error
that arise when mapping smallholder croplands. To overcome errors caused
by mismatches between image resolution and cropland scales, as well as
persistent cloud cover, the system converts daily, 3.7 m PlanetScope
imagery into two seasonal composites within a single agricultural year.
To reduce errors that occur when training classifiers, we built a
labelling platform that rigorously assesses label accuracy, and creates
more accurate consensus labels that train a Random Forests model. The
labelling platform and model interact within an active learning process
that boosts the accuracy of the resulting cropland probability map,
which is used in a segmentation process to delineate individual field
boundaries. We applied this system to map Ghana's croplands for the year
2018. We divided Ghana into 16 mapping regions (12,160-23,535 km\(^2\)),
training separate models for each using a total of 6,299 labels, plus
1,600 for validation. Using an independent map reference sample
(n=1,207), we found that overall accuracies of the resulting cropland
probability and field boundary maps were 88\% and 86.7\%, respectively,
with User's accuracies for the cropland class of 61.2\% and 78.9\%, and
Producer's accuracies of 67.3\% and 58.2\%. Croplands covered
16.1-23.2\% of the mapped area, comprising 1,131,146 total fields with
an average size of 3.92 ha. Estimates based on the map reference sample
indicate the cropland percentage is 17.1\% (15.4-18.9\%) or 17.6\%
(15.6-19.6\%), depending on the map used to estimate the standard error.
Using the labellers' digitized field boundaries to estimate biases in
field boundary statistics, we calculated an adjusted mean field size of
1.73 ha and total field count of 1,662,281. Although the cropland class
contained substantial errors, the system was effective in mitigating
error and quantifying resulting performance gains. By minimizing
training errors, consensus labelling improved the model's F1 scores by
up to 25\%, while 3 iterations of active learning increased the F1 score
by 9.1\%, on average, which was 2.3\% higher than training models with
randomly selected labels. Map accuracy can be improved by replacing
Random Forests with a convolutional neural network. These results
demonstrate a readily adapted, transferrable framework for developing
high resolution, annual, nation-scale maps that provide important
details about smallholder-dominated croplands.
\end{abstract}
\end{singlespace}

@@

Amidst all the challenges posed by global change, a particular concern
is how agricultural systems will adapt to meet humanity's growing food
demands, and the impacts that transforming and expanding food systems
will have on societies, economies, and the environment (Searchinger et
al. 2019). Significant efforts are being made to address the various
aspects of this challenge, including work on diagnosing and closing
yield gaps (Lobell et al. 2009, e.g. Licker et al. 2010, Mueller et al.
2012), expanding and commercializing production (Morris and Byerlee
2009), and to understand (Rulli and D'Odorico 2014, Kehoe et al. 2017,
Davis et al. 2020) and mitigate (Estes et al. 2016b) agriculture's
ecological impacts. Answering many of the questions these efforts seek
to address depends on reliable data that describes the location and
characteristics of cropland (Fritz et al. 2015), and how these are
changing over time, ideally on an annual basis, given the rapid pace of
change (Gibbs et al. 2010, Bullock et al. 2021). Unfortunately, the data
that do exist are in many places inaccurate and infrequently updated.
Existing estimates of how much global cropland there is tend to vary
widely, and they often disagree about where cropland is located (e.g.
Fritz et al. 2011, 2013). Such errors can propagate in subsequent
analyses that rely on cropland data as inputs, resulting in potentially
misleading answers (Estes et al. 2018). Beyond cropland distributions,
few data are available on key cropland characteristics such as field
size, an important variable needed to estimate yield and other key food
security variables (Carletto et al. 2015), and as an indicator of farm
size (Levin 2006, Samberg et al. 2016), a critical component of rural
livelihoods given increasing population densities and longstanding
debates about the relationship between farm size and productivity (Feder
1985, Carletto et al. 2013, Desiere and Jolliffe 2018).

These informational inadequacies are due to the fact that cropland data
in much of the world are derived from remotely sensed landcover maps,
which can be notoriously high in error, particularly over regions such
as Africa (Fritz et al. 2010, Estes et al. 2018), where agricultural
changes will be largest and the need for accurate baseline data is thus
greatest (Searchinger et al. 2015, Estes et al. 2016b, Bullock et al.
2021). Cropland mapping over Africa is difficult for several reasons.
The primary reason relates to the characteristics of the continent's
smallholder-dominated croplands, where half of all fields are smaller
than 1 ha (Lesiv et al. 2019). This size is small relative to the 30-250
m resolution of the sensors typically used in many landcover mapping
efforts (e.g. Chen et al. 2015, Sulla-Menashe et al. 2019), which
results in errors due to mixed pixels and aspects of the modifiable area
unit problem (Openshaw and Taylor 1979, Boschetti et al. 2004){]}. In
the latter case, the pixel's shape may be poorly matched to that of
cropland, and is too coarse to aggregate to approximate that shape at
the characteristic scales of crop fields (Dark and Bram 2007, Estes et
al. 2018). On top of the matter of scale is 1) high intra-class
variability of the cropland class, compounded by the fact that these
particular croplands can be heavily intergraded with surrounding
vegetation (Debats et al. 2016, Estes et al. 2016a), and 2) the
substantial temporal variability within croplands, both within and
between seasons. These latter two aspects pose challenges for the
classification algorithms that are applied to the imagery.

Recent technological advances are helping to overcome these challenges.
Chief among these are the growing numbers of satellites that collect
high (\textless5 m) to near-high (10 m) resolution imagery at sub-weekly
intervals (Drusch et al. 2012, McCabe et al. 2017). This high spatial
\emph{and} temporal resolution imagery addresses the sensor-field scale
mismatch, and more effectively captures the intra-seasonal dynamics of
cropland, which helps classifiers distinguish cropland from surrounding
cover types (Debats et al. 2016, Defourny et al. 2019). On top of this,
advances in cloud computing and the opening of satellite image archives
(Wulder et al. 2016), and next generation machine learning approaches
are placing large volumes of these moderate to near-high resolution
imagery together with the computational and algorithmic resources
necessary to classify them at scale (Gorelick et al. 2017). These
capabilities are already being used to create a new generation of higher
resolution (10-30 m) cropland and landcover maps for Africa and other
regions {[}ESA (n.d.); Lesiv et al. (2017); Xiong et al. (2017); (Zhang
et al. 2021){]}. However, the potential of the highest resolution
(\textless5 m) imagery to map cropland over very large extents
(e.g.~country scales) has yet to be realized, presumably because these
data are commercial and relatively expensive, and require significant
computational resource to process.

Beyond the imagery and computational gains, machine learning algorithms
are rapidly advancing, providing large gains in classification
performance (Maxwell et al. 2018, Ma et al. 2019). However, the ability
to take advantage of these gains is often limited by newer models' need
for large training datasets, which are typically unavailable, hard to
collect, or contain numerous errors in developing regions (Ma et al.
2019, Elmes et al. 2020, Burke et al. 2021). To build sufficient samples
for training cropland mapping models, as well as the reference data
needed to objectively assess their performance (we refer collectively to
both sample types as ``labels,'' distinguishing between each type where
needed), map-makers rely heavily on visual interpretation of high
resolution satellite or aerial imagery (Chen et al. 2015, e.g. Xiong et
al. 2017, Stehman and Foody 2019), as it is impractical and expensive to
collect these data in the field over large areas, particularly on an
ongoing basis. Several web-based platforms have been developed to
facilitate such efforts, which provide convenient and highly scalable
tools for training data collection (Fritz et al. 2012, Estes et al.
2016a, e.g. Bey et al. 2016). Visually interpreted labels present two
particular problems. The first is that they inevitably contain errors of
interpretation that can vary substantially according to the skill of the
labeller, particularly over complex croplands with small field sizes
(Estes et al. 2016a, Waldner et al. 2019). The second problem is that
visual interpretation depends on high resolution imagery (\textless5 m),
as field boundaries become increasingly difficult to discern as image
resolution decreases. Typically the only practical source of high
resolution imagery is ``virtual globe'' basemaps (e.g.~Bing or Google
Maps), which are composed of mosaics of various high resolution
satellite and aerial images that typically span 3-5 years of time within
a single country (Lesiv et al. 2018). This within-mosaic temporal
variation can set up a temporal mismatch between the imagery being
interpreted and the imagery being classified, which is usually from a
different source (e.g.~Landsat, Sentinel; Xiong et al. (2017)). If a
land change occurs in the interval between the two image sets (e.g.~a
new field was created), the label, even if accurately drawn, introduces
error into the classifier. This source of error may be elevated in
tropical smallholder systems, where swidden agricultural is common (Van
Vliet et al. 2013), or in rapidly developing agricultural frontiers
(Zeng et al. 2018). Despite the high potential for label error, they are
typically not accounted for during model training and map accuracy
assessment, resulting not only in the potential for maps to be misused
or misinterpreted in subsequent analyses, but in missed opportunities to
improve model performance (Estes et al. 2018, Stehman and Foody 2019,
Elmes et al. 2020).

Taking into consideration the advances and remaining limitations
described above, the ability to map smallholder-dominated croplands can
be further improved by 1) more fully exploiting the profusion of high
frequency, high resolution imagery provided by CubeSats (McCabe et al.
2017), and 2) by implementing methods that improve the ability to
collect and minimize errors in image-interpreted labels. We developed a
mapping approach that focuses on these two sources of improvement. Our
approach uses PlanetScope imagery collected by Planet's fleet of Dove
satellite, which provides 3-4 m resolution imagery over large areas at
near daily intervals (McCabe et al. 2017, PlanetTeam 2018), at
relatively low to no cost for academic
research\footnote{www.planet.com/markets/education-and-research/} and
non-commercial, sustainability-oriented
applications\footnote{assets.planet.com/docs/Planet\_ParticipantLicenseAgreement\_NICFI.pdf}.
Although these data are of lower spectral depth and, in some cases,
quality, than Landsat, Sentinel, or Worldview imagery, their daily
revisit enables country- to continent-scale image mosaics to be created
for multiple periods during a single agricultural year, even over the
cloudiest forest regions where it is hard to successfully construct
cloud-free composites from optical imagery with return intervals (even
by a few days). This ability to capture intra-annual variability can be
more important for classifying cropland than spectral depth (Debats et
al. 2016). Beyond the frequency, PlanetScope's 3.7 m
resolution--although substantially coarser the 0.5-1 m available in most
areas covered by virtual globes--is sufficiently resolved for humans to
discern small fields under many conditions (Fourie 2009, e.g.~see Estes
et al. 2018). This allows labels to be made using the same imagery that
is classified, which helps to minimize label error. To further reduce
label noise, we developed a platform that includes rigorous label
accuracy assessment protocols and a novel approach for creating
consensus labels, which helps reduce mistakes made by individual
labellers (Estes et al. 2016a, Elmes et al. 2020). We couple the
labelling platform with a machine learning model inside an active
learning (Cohn et al. 1994, Tuia et al. 2011) framework, in which model
training is done iteratively and interactively, wherein the model's
prediction uncertainty over unlabelled areas is assessed after each
training step, and a sample of the most uncertain sites is selected for
additional labelling (Cohn et al. 1994, Tuia et al. 2011). This approach
helps boost the performance of the classifier while reducing the overall
number of labels required to achieve a given level of performance
(Debats et al. 2017, e.g. Hamrouni et al. 2021). Finally, an
unsupervised segmentation step is applied to convert pixel-wise cropland
predictions into vectorized maps of individual field boundaries.

Here we use this approach to create a high resolution, country-scale map
of crop field boundaries in Ghana, a country where smallholder farming
predominates, and which has a broad mix of climates and agricultural
systems, including large areas where shifting agriculture is practiced
(Samberg et al. 2016, Kansanga et al. 2019). The map we develop
represents a single agricultural year (2018-2019), as opposed to a
multi-year composite, thereby demonstrating a capacity for annual, high
resolution maps that can be used to monitor rapidly evolving small-scale
agricultural systems, including key characteristics such as field size.
In addition to presenting one of the most spatially extensive
agricultural applications of CubeSats to date, we provide a new
technique for converting daily imagery into seasonal composites, and
show how best practices for model training and label collection can be
applied to improve map accuracy (Elmes et al. 2020).

@@

The cropland mapping approach we developed is comprised of four primary
components (Figure \ref{fig:systemoverview}) that were developed using
open source software, and is designed to run in a cloud computing
environment. The first component collects daily PlanetScope imagery for
a period of several months over a particular mapping geography, and
converts these into cloud-free seasonal composites. The second component
is a custom-built platform that a provides a set tools for labelling the
PlanetScope composites, and built-in procedures for assessing and
minimizing label error. This platform interacts with the third
component, a machine learning process running in Apache Spark, which
includes a routine for 1) selecting new training sites based on their
prediction uncertainty, and 2) returning these to the labelling platform
for labeling. This interaction, which represents active learning (Cohn
et al. 1994, Tuia et al. 2011), repeats until a model performance
threshold is reached. The fourth and final component is a segmentation
process that is applied to the seasonal image composites to segment the
entire image, and then merged with the pixel-wise probability maps
generated by active learning to identify the segments corresponding to
croplands, resulting in a final vectorized field boundaries. Each system
component is available on an open repository (see data and software
availability section for details).

@@

\caption{An overview of the primary system components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them. The dashed line indicates iterative interactions, while solid lines indicate one-time or irregular connections.}\label{fig:systemoverview}
\end{figure}

@@

The image processing components of our system were designed to work with
PlanetScope Analytic surface reflectance data (PlanetTeam 2018).
PlanetScope provides three visual (red, green, blue) and a near-infrared
band at 3-4 m resolution at nominal daily frequency. Although these
images are already pre-preprocessed and corrected for atmospheric
effects, there are residual errors from inter-sensor differences and the
radiometric normalization process (Houborg and McCabe 2018), variation
in the orientation of scene footprints, as well as a high frequency of
cloud cover over the study region (Wilson and Jetz 2016). To correct for
these factors, we developed a procedure for creating temporal composites
representing the primary growing and non-growing seasons within a single
year.

PlanetScope imagery is accessed via the Planet API (PlanetTeam 2018),
and an initial order is placed for all imagery falling within the
mapping geography and the date ranges for the two compositing periods.
The imagery is collected and transferred directly to a cloud storage
platform (Amazon Web Services {[}AWS{]} S3).

Individual images are then transformed into analysis ready data (ARD)
(Dwyer et al. 2018), by subsetting each downloaded image into 0.05
degree tiles, regardless of cloud cover. Tiles are organized within a
larger 1 degree resolution grid that covers the entire continent, which
defines the minimum mapping area of interest (AOI; Figure
\ref{fig:aois}).

{\centering \includegraphics[width=1\linewidth]{figures/figure2}

\caption{The reference system used in the mapping platform, including A) numbered areas of interest (AOIs) that define the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), B) the 0.05 degree tile used for compositing imagery, and C) the 0.005 degree resolution reference grid used for collecting training data and distributed computing.}\label{fig:aois}
\end{figure}

The temporal compositing process is applied to the tiled daily images
for the time period of interest, which in this case was one of two
multi-month seasons, the primary growing and dry seasons for a single
agricultural year. Imagery from two seasons helps to improve the
performance of cropland classifiers (Debats et al. 2016), while having
the seasons in the same year helps to minimize differences caused by
land change. For each pixel in each image in each ARD temporal stack for
a given season, two weights are calculated:

@@

Where \emph{t} is a particular date in near-daily time series of
PlanetScope images, which begins at date 1 for the given compositing
period and ends on date \emph{i}, \emph{blue} is the blue band, and
\emph{NIR} the near infrared band. Equation \ref{eq:cloud} assigns lower
weights to hazy and clouded pixels as the blue band is sensitive to haze
and cloud pixels (Zhang et al. 2002), while Equation \ref{eq:shadow}
assigns low weights to pixels in cloud shadow considering the
significant darkening effect of the cloud shadows in the Near Infrared
band (Zhu and Woodcock 2012, Qiu et al. 2020)

Once these two weights are calculated, the final composited pixel value
for each of the four PlanetScope bands is:

@@

Which is a weighted mean for each pixel for each band \emph{B} for the
particular compositing period. The composited tiles were then added to
the S3 store (Figure \ref{fig:systemoverview}), where they are stored as
cloud-optimized geotiffs, and a ``slippy
map\footnote{https://wiki.openstreetmap.org/wiki/Slippy\_Map}''
rendering is created for each composite using Raster Foundry (Azavea
2020). The web-rendered imagery is presented within the training data
platform (next section).

\hypertarget{labelling-platform}{%
\subsubsection{Labelling platform}\label{labelling-platform}}

Training and reference data are collected by a custom labelling
platform, which was originally designed for AWS's Mechanical Turk job
marketplace (Estes et al. 2016a). The basic structure of the system
remains the same, but we converted it into a standalone platform that
allows us to enroll and pay people directly for their labelling, and is
designed to control and supervise the machine learning process. The
platform runs on a Linux virtual machine hosted on an AWS EC2 instance
and is comprised of a database (PostGIS/Postgres), a mapping interface
(OpenLayers 3), an image server (Raster Foundry), and a set of utilities
for managing, assessing, and converting digitization work into
rasterized labels for training a machine learning algorithm. Each
instance of the platform focuses on a specific AOI (Figure
\ref{fig:aois}A)

The following sections provide an overview of the labelling platform's
architecture.

\hypertarget{mapping-workflow}{%
\paragraph{Mapping workflow}\label{mapping-workflow}}

\hypertarget{selecting-training-and-reference-sites}{%
\subparagraph{Selecting training and reference
sites}\label{selecting-training-and-reference-sites}}

The labelling process begins with the random selection of a subset
(e.g.~500) of cells from a 0.005 degree grid, with the selection itself
potentially split into a training and validation sample, according to
predetermined proportions. The grid, which is nested within the tiling
and larger 1 degree grids (Figure \ref{fig:aois}C) defines the spatial
unit for a labelling job. The selected cells are placed into a queue
within the platform's database, and then converted into a mapping
\emph{task} that has a specified number of \emph{assignments}
(boundaries drawn by an individual labeller) that must be completed
before the task is complete.

\hypertarget{mapping-assignments}{%
\subparagraph{Mapping assignments}\label{mapping-assignments}}

>>>>

%\begin{center}
This pre-print not yet undergone peer review. It has been submitted to
\emph{Frontiers in Artificial Intelligence}. This version will be
updated as it is revised, and the final published version will be
accessible through its DOI link.
%\end{center}

@@

\begin{abstract}
Mapping the changing characteristics of Africa's smallholder-dominated
agricultural systems, including the sizes and numbers of fields, is
crucial for understanding food security and a range of other
socioeconomic and environmental concerns. However, accurately mapping
these systems is difficult because of 1) the spatial and temporal
mismatch between satellite sensors and smallholder fields, and 2) the
lack of high-quality labels needed to train and assess machine learning
classifiers. We developed an approach designed to address these two
problems, which we used to map Ghana's annual croplands for the year
2018. To overcome the first problem, we converted daily, high resolution
CubeSat (PlanetScope) imagery into two cloud-free seasonal composites
covering a single agricultural year. To address the second problem, we
created a labelling platform that rigorously assesses and minimizes
label error, and used it to iteratively train a Random Forests
classifier with active learning, which identifies the most informative
training sample based on prediction uncertainty. Minimizing label errors
improved model F1 scores by up to 25\%. Active learning increased F1
scores by an average of 9.1\% between first and last training
iterations, and 2.3\% more than models trained with randomly selected
labels. We used the resulting 3.7 m map of cropland probabilities within
a segmentation algorithm to delineate crop field boundaries. Based on an
independent map reference sample (n=1,207), the cropland probability and
field boundary maps have respective overall accuracies of 88\% and
86.7\%, user's accuracies for the cropland class of 61.2\% and 78.9\%,
and producer's accuracies of 67.3\% and 58.2\%. Using the map reference
sample to calculate an unbiased area estimate from the field boundary
map, we found that cropland covers 17.1\% (15.4-18.9\%) of Ghana. Using
the most accurately digitized labels to calculate and correct for biases
in the segmented field boundaries map, we further estimated the average
size (1.73 ha) and total number (1,662,281) of crop fields in Ghana. Our
results demonstrate an adaptable and transferrable approach for mapping
the characteristics of croplands on an annual basis and over national
extents, with several features that effectively mitigate the errors
inherent in remote sensing of smallholder-dominated agriculture.
\end{abstract}
\end{singlespace}

@@

Amidst all the challenges posed by global change, a particular concern
is how agricultural systems will adapt to meet humanity's growing food
demands, and the impacts that transforming and expanding food systems
will have on societies, economies, and the environment (Searchinger et
al. 2019). A number of efforts are underway to address various aspects
of this challenge, including work on diagnosing and closing yield gaps
(Lobell et al. 2009, e.g. Licker et al. 2010, Mueller et al. 2012),
expanding and commercializing production (Morris and Byerlee 2009), and
to understand (Rulli and D'Odorico 2014, Kehoe et al. 2017, Davis et al.
2020) and mitigate (Estes et al. 2016b) agriculture's ecological
impacts. The success of these efforts depends heavily on data that
accurately describes the location and characteristics of croplands
(Fritz et al. 2015), and, given the rapid pace of agricultural change
(Gibbs et al. 2010, Zeng et al. 2018, Bullock et al. 2021), how these
are changing from one year to the next. Unfortunately, for many regions,
existing cropland datasets are inaccurate, and are usually created as
once-off or infrequently update products. As such, estimates of global
cropland area tend to vary widely, often disagree about where croplands
are located (e.g. Fritz et al. 2011, 2013), and become rapidly outdated.
Errors in these maps can propagate in subsequent analyses that use
cropland data as inputs, resulting in potentially misleading answers
(Estes et al. 2018). Beyond distributions, few data are available on key
cropland characteristics such as field size, an important variable
needed to estimate yield and other key food security variables (Carletto
et al. 2015), and as an indicator of farm size (Levin 2006, Samberg et
al. 2016), a critical component of rural livelihoods given increasing
population densities and longstanding debates about the relationship
between farm size and productivity (Feder 1985, Carletto et al. 2013,
Desiere and Jolliffe 2018).

The deficit of information is due to the fact that in many regions the
only source of cropland data are remotely sensed land cover maps, which
are prone to error. This is particularly true in Africa (Fritz et al.
2010, Estes et al. 2018), where agricultural changes will be largest and
the need for accurate baseline data is thus greatest (Searchinger et al.
2015, Estes et al. 2016b, Bullock et al. 2021), and where the
characteristics of croplands exacerbate the error inherent in remote
sensing analyses. Half of all fields in Africa's smallholder-dominated
agricultural systems are smaller than 1 ha (Lesiv et al. 2019). This
size is small relative to the 30-250 m resolution of the sensors
typically used in many landcover mapping efforts (e.g. Chen et al. 2015,
Sulla-Menashe et al. 2019), which results in errors due to mixed pixels
and aspects of the modifiable area unit problem (Openshaw and Taylor
1979, Boschetti et al. 2004), wherein the pixel's shape does not match
that of crop fields, and is too coarse to aggregate into an
approximation of that shape (Dark and Bram 2007, Estes et al. 2018). On
top of the matter of scale is the high variability within and between
fields, their tendency to intergrade with surrounding vegetation (Debats
et al. 2016, Estes et al. 2016a), and the high temporal variability
within croplands. These last three aspects pose challenges for the
classification algorithms that are applied to the imagery.

Recent technological advances are helping to overcome these challenges.
Chief among these are the growing numbers of satellites that collect
high (\textless5 m) to near-high (10 m) resolution imagery at sub-weekly
intervals (Drusch et al. 2012, McCabe et al. 2017). The spatial
resolution of these imagery addresses the scale mismatch between sensor
and field, and their high frequency captures the seasonal dynamics of
cropland, which helps classifiers distinguish cropland from surrounding
cover types (Debats et al. 2016, Defourny et al. 2019). On top of this,
the opening of satellite image archives (Wulder et al. 2016) and
advances in cloud computing are placing large volumes of moderate to
near-high resolution imagery together with the computational and
algorithmic resources necessary to classify them at scale (Gorelick et
al. 2017). These capabilities have already been used to create a new
generation of higher resolution (10-30 m) cropland and landcover maps
for Africa and other regions {[}ESA (n.d.); Lesiv et al. (2017); Xiong
et al. (2017); (Zhang et al. 2021){]}. However, the potential of the
highest resolution (\textless5 m) imagery to map cropland over very
large extents (e.g.~country scales) has yet to be realized, presumably
because these data are commercial and relatively expensive, and require
significant computational resource to process.

Beyond the imagery and computational gains, machine learning algorithms
are rapidly advancing, providing large gains in classification
performance (Maxwell et al. 2018, Ma et al. 2019). However, the ability
to take advantage of these gains is often limited by newer models' need
for large training datasets, which are typically unavailable, hard to
collect, or contain numerous errors (Ma et al. 2019, Elmes et al. 2020,
Burke et al. 2021). To build sufficient training samples, as well as the
reference data needed to objectively assess their performance (we refer
collectively to both types as ``labels,'' distinguishing between each as
needed), map-makers rely heavily on visual interpretation of high
resolution satellite or aerial imagery (Chen et al. 2015, e.g. Xiong et
al. 2017, Stehman and Foody 2019), as it is impractical and expensive to
collect these data in the field over large areas, particularly on an
ongoing basis. Consequently, a number of web-based platforms have been
developed to collect such labels (Fritz et al. 2012, Estes et al. 2016a,
e.g. Bey et al. 2016). Image-drawn labels present two particular
problems. The first is that they inevitably contain errors of
interpretation, which can vary substantially according to the skill of
the labeller, particularly over complex croplands with small field sizes
(Estes et al. 2016a, Waldner et al. 2019). The second problem is that
visual interpretation depends on high resolution imagery (\textless5 m),
as fields are increasingly difficult to discern as image resolution
decreases. Typically the only available source of high resolution
imagery is ``virtual globe'' basemaps (e.g.~Bing or Google Maps), which
present mosaics of high resolution satellite and aerial images collected
over a span of several years (Lesiv et al. 2018). This within-mosaic
temporal variation can create a temporal mismatch between the labels and
the imagery being classified, which is usually from a different source
(e.g.~Landsat, Sentinel; Xiong et al. (2017)). If a land change occurs
in the interval between the two image sets (e.g.~a new field was
created), the label, even if accurately drawn, introduces error into the
classifier. This source of error may be elevated in croplands where
swidden agriculture is practiced (Van Vliet et al. 2013), or in rapidly
developing agricultural frontiers (Zeng et al. 2018). Despite the high
potential for it, label error is often not considered during model
training and map accuracy assessment, resulting not only in the
potential for maps to be misused or misinterpreted, but in missed
opportunities to improve model performance (Estes et al. 2018, Stehman
and Foody 2019, Elmes et al. 2020).

Taking into consideration the advances and remaining limitations
described above, the ability to map smallholder-dominated croplands can
be further improved by 1) more fully exploiting the profusion of high
frequency, high resolution imagery provided by CubeSats (McCabe et al.
2017), and 2) by implementing methods that improve the ability to
collect and minimize errors in image-interpreted labels. We developed a
mapping approach that focuses on these two sources of improvement. Our
approach uses PlanetScope imagery collected by Planet's fleet of Dove
satellite, which provides 3-4 m resolution imagery over large areas at
near daily intervals (McCabe et al. 2017, PlanetTeam 2018), at
relatively low to no cost for academic
research\footnote{www.planet.com/markets/education-and-research/} and
non-commercial, sustainability-oriented
applications\footnote{assets.planet.com/docs/Planet\_ParticipantLicenseAgreement\_NICFI.pdf}.
Although these data are of lower spectral depth and, in some cases,
quality, than Landsat, Sentinel, or Worldview imagery, their daily
revisit enables country- to continent-scale image mosaics to be created
for multiple periods during a single agricultural year, even over the
cloudiest forest regions where it is hard to successfully construct
cloud-free composites from optical imagery with return intervals (even
by a few days). This ability to capture intra-annual variability can be
more important for classifying cropland than spectral depth (Debats et
al. 2016). Beyond the frequency, PlanetScope's 3.7 m
resolution--although substantially coarser than the 0.5-1 m imagery
available in most areas covered by virtual globes--is sufficiently
resolved for humans to discern small fields under many conditions
(Fourie 2009, e.g.~see Estes et al. 2018). This allows labels to be made
using the same imagery that is classified, which helps to minimize label
error. To further reduce label noise, we developed a platform that
includes rigorous label accuracy assessment protocols and a novel
approach for creating consensus labels, which helps reduce mistakes made
by individual labellers (Estes et al. 2016a, Elmes et al. 2020). We
couple the labelling platform with a machine learning model inside an
active learning (Cohn et al. 1994, Tuia et al. 2011) framework, in which
the model is trained interactively, using the model's prediction
uncertainty over unlabelled areas to select new sites for additional
labelling (Cohn et al. 1994, Tuia et al. 2011). This approach helps
boost the performance of the classifier while reducing the overall
number of labels required to achieve a given level of performance
(Debats et al. 2017, e.g. Hamrouni et al. 2021). An unsupervised
segmentation step is then applied to convert pixel-wise cropland
predictions into vectorized maps of individual field boundaries.

Here we use this approach to create a high resolution, country-scale map
of crop field boundaries in Ghana, a country where smallholder farming
predominates across a broad mix of climate and agricultural systems,
ranging from primarily grain and vegetable crop production in the
northern savannas to tree crop-dominated systems in the forested
southwest, including large areas where shifting agriculture is practiced
(Samberg et al. 2016, Kansanga et al. 2019). The map represents a single
agricultural year (2018-2019), as opposed to a multi-year epoch, thereby
demonstrating a capacity for annual, high resolution maps that can be
used to monitor rapidly evolving small-scale agricultural systems,
including key characteristics such as field size. In addition to
providing valuable new data and insight into Ghana's agriculture, our
study demonstrates one of the most spatially extensive agricultural
applications of CubeSats to date, provides a new technique for
converting daily imagery into seasonal composites, and shows how best
practices for model training and label collection can be applied to
improve map accuracy (Elmes et al. 2020).

@@

The mapping approach we developed is comprised of four open source
components (Figure \ref{fig:systemoverview}) that are designed to run in
a cloud computing environment. The first component collects daily
PlanetScope imagery and converts them into cloud-free seasonal
composites. The second is a custom-built platform that provides tools
for labelling the composites, along with procedures to assess and
minimize label error. This platform interacts with the third component,
a machine learning process, within an active learning (Cohn et al. 1994,
Tuia et al. 2011) loop, to produce a map of predicted cropland
probabilities for each image pixel. The fourth and final component is an
algorithm that segments the image composites, then filters the resulting
polygons using the pixel-wise cropland predictions produced by the
active learning classifier, resulting in a final set of vectorized field
boundaries.

@@

\caption{An overview of the primary mapping components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them. The dashed line indicates iterative interactions, while solid lines indicate one-time or irregular connections.}\label{fig:systemoverview}
\end{figure}

We describe each component in further detail in the following section,
and how we applied them to map Ghana's annual cropland boundaries,
excluding tree crops.

@@

The image processing component was designed for PlanetScope Analytic
surface reflectance imagery (PlanetTeam 2018), which provides three
visual (red, green, blue) and near-infrared bands at 3.7 m resolution at
nominal daily frequency. The images are provided as ortho-rectified and
converted to surface reflectance, although there are residual errors
from inter-sensor differences and the radiometric normalization process
(Houborg and McCabe 2018), variation in the orientation of scene
footprints, as well as a high frequency of cloud cover over the study
region (Wilson and Jetz 2016, Roy et al. 2021) that are not fully
captured by the provided cloud masks. To minimize the effect of these
residual errors, we developed a procedure for creating temporal
composites of the primary growing and non-growing seasons within a
single 12-month period. For Ghana, we defined the primary growing season
as May through September, followed by the off (or dry) season from
November or December through February. We chose these two seasons
because prior work shows that the contrast between them improves
cropland classifications (Debats et al. 2016), Furthermore, capturing
the seasons in this sequence during the same year helps minimize
differences caused by land change. The wide time intervals we used to
define each season were necessary for collecting a sufficient number of
images to make high quality composites, as Ghana's cloud cover renders
many scenes unusable and therefore unavailable in Planet's catalog, thus
the effective return interval can be substantially longer than 24 hours
during the cloudiest months (Roy et al. 2021).

We collected all available scenes intersecting Ghana and falling within
these two seasons during the 2018 agricultural year (defined here as
March, 2018-February, 2019) via the Planet API (PlanetTeam 2018), and
transferred these to cloud storage (Amazon Web Services {[}AWS{]} S3).
We then converted each scene into analysis ready data (Dwyer et al.
2018) by cropping each to the boundaries of a 0.05\(^\circ\) grid that
it intersected (see Figure S1 in Supplemental Information {[}SI{]}),
which provided the dimensions for making composited image tiles. We
chose this cell size for tiling because it is slightly narrower than the
short axis of a PlanetScope scene, which increases the number of
intersecting scenes that completely cover the tile, thereby helping to
minimize edge artifacts in the composites.

To create a seasonal composite, we calculated two weights for the time
series of each pixel within the ARD stack for a given season:

@@

Where \emph{t} is a particular date in the pixel time series, which
begins at date 1 for the given compositing period and ends on date
\emph{i}, \emph{blue} is the blue band, and \emph{NIR} the near infrared
band. Equation \ref{eq:cloud} assigns lower weights to hazy and clouded
pixels as the blue band is sensitive to these atmospheric features
(Zhang et al. 2002), while Equation \ref{eq:shadow} assigns low weights
to pixels in cloud shadow (Zhu and Woodcock 2012, Qiu et al. 2020)

After assigning these two weights, we calculated the final composited
pixel value:

@@

Which is the weighted mean for each pixel for each band \emph{B} for the
given season.

Each composited seasonal tile was saved as a cloud-optimized geotiff,
and a ``slippy
map\footnote{https://wiki.openstreetmap.org/wiki/Slippy\_Map}''
rendering was created for each composite using Raster Foundry (Azavea
2020), for display within the labelling platform (next section).

We generated a catalog of 16232 composite tiles (hereafter simply
``tiles'') for Ghana, consisting of a seasonal pair for each of the 8116
0.05\(^\circ\) tile grid cells covering Ghana. To assess the quality of
the resulting composites, 50 tile grid cells were randomly selected, and
two separate observers graded each corresponding seasonal composite
using a four category that evaluated the degree of 1) residual cloud and
2) cloud shadow, 3) the number of visible scene boundary artifacts, and
4) the proportion of the image with resolution degraded below the 3.7 m
PlanetScope resolution (e.g.~because of between-date image
mis-registrations). Each category was qualitatively ranked from 0-3,
with 0 being the lowest quality, and 3 the highest (see SI for complete
protocol), making the highest possible score 12. We rescaled scores to
fall between 0 and 1.

\hypertarget{mapping-cropland-probabilities-with-active-learning}{%
\subsection{Mapping cropland probabilities with active
learning}\label{mapping-cropland-probabilities-with-active-learning}}

The first step in creating a country-wide field boundary map of Ghana
was to create a pixel-wise classification of cropland probabilities
throughout the country. Given the high resolution of the imagery and the
need to minimize the computational burden, we divided Ghana into 16
distinct mapping regions, or Areas of Interest (AOIs). We constructed
the AOIs by grouping together tile grids into blocks representing the
larger 1\(^\circ\) cells used to assign tile identifiers (Figure S1A).
We grouped tile cells from 1\(^\circ\) degree cells that overlapped
Ghana's boundaries together with those from the nearest 1\(^\circ\) cell
contained entirely within Ghana (with the exception of AOI 16, which was
comprised of tile grids from the 1\(^\circ\) cells along Ghana's
southern coast. The average extent of the resulting AOIs was 15,457
km\(^2\) (range 12,160-23,535 km\(^2\)).
